Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}

### 2017-08-17 ident_manu.tex
almost done with revision

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Sun Aug 6 08:21:57 2017 -0500, \href{https://github.com/fawda123/identifiability/commit/c03d0df749cea19daced8a68105265cf1ed2a63e}{c03d0df749cea19daced8a68105265cf1ed2a63e}\newline Acronyms: \acfi{chla}; \acfi{cgem}; \acfi{do}; \acfi{dom}; \acfi{gom}; \acfi{lcs}; \acfi{marb}; \acfi{pom}; \acfi{rmse}; \acfi{zerod}}

@@

\begin{abstract}
\noindent Local sensitivity analyses and identifiable parameter subsets were used to describe numerical constraints of a hypoxia model for bottom waters of the northern Gulf of Mexico.  The sensitivity of state variables differed considerably with parameter changes, although most variables were responsive to changes in parameters that influenced planktonic growth rates and less sensitive to physical or chemical parameters.  Variation in sensitivity had a direct correspondence with identifiability, such that only small subsets of the complete parameter set had unique effects on the model output. Selecting parameters by decreasing sensitivity demonstrated that only eight of 51 total parameters had a sufficiently unique effect on model output for accurate calibration.  As a result, parameter selection heuristics were used to identify parameters for model calibration that depended on combined effects on output, relative sensitivity of each parameter, and ecological categories for the biogeochemical equations. The calibrated \ac{zerod} unit of the hypoxia model had improved fit to the observed data if sensitive parameters were included in an identifiable subset.  Extension of results to a three-dimensional grid of the Gulf of Mexico showed that sensitive parameters for the \ac{zerod} model translated to non-trivial changes in the areal estimates of hypoxia.
\end{abstract}

@@

Selected parameter subsets from the seven heuristics were used to calibrate the \ac{zerod} model.  This analysis demonstrated that different parameter sets can produce model output with differing goodness and fit and calibration effort given
\begin{inparaenum}[1\upshape)]
\item differences in variable sensitivity to parameters in each subset, 
\item differences in the parameters selected from different categories (heuristic difference), and 
\item differences in identifiability values.
\end{inparaenum}
Generally, this analysis was a proof of concept that the selection heuristics can lead to an `optimal' model by maximizing the balance between sensitivity and identifiability.

A time series of \ac{do} from Weeks Bay, Alabama was used for model calibration.  Daily observations from January 1\textsuperscript{st},  2006 to January 1\textsuperscript{st}, 2007 at the Weeks Bay station of the Weeks Bay National Estuarine Reserve were used \citep{Wenner04}.  We chose this location because of the availability of data for model calibration compared to bottom waters in the northern \ac{gom} and the system exhibits hypoxia from eutrophication.  This application also provided a practical example of how the FishTank model could be calibrated using data from alternative locations.  Initial conditions for model input were modified to match conditions in Weeks Bay using observed water quality and nutrient data (all data from \url{http://cdmo.baruch.sc.edu/}).  The \ac{do} time series was filtered to reduce physical effects of the tidal advection component on the observed data (\textit{sensu} \citealt{Beck15b}).  As such, the filtered data were comparable to a time series with no physical advection to match conditions of the \ac{zerod} FishTank model with no hydrodynamic component.

Each parameter subset was calibrated to most closely match the model output to the observed time series.  Parameter values were searched using the optim function in R that implemented the limited memory modification of the BFGS quasi-Newton method \citep{Byrd95,Nocedal06,RDCT17}.  This optimization scheme searches for values within the user-defined lower and upper bounds of each parameter.  We set the algorithm to search within +/- 50\% of the default parameter values.  While we recognize that `actual' parameter values could possibly extend beyond this range, constraining the values to the range tested herein was expected to produce output that was more interpretable within the context of the sensitivity functions defined above.  Moreover, precise limits on each parameter were less of a concern for the current analysis as our primary goal was to demonstrate differences in fit as a function of sensitivity and identifiability.  The optimal values for each parameter were identified based on convergence of the model results as a minimization of the \ac{rmse} between the observed and modelled data.

@@

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in 38 of the 51 (75\% of total) parameters that were evaluated in FishTank (\cref{fig:sensalltile}, \cref{tab:dosens}). Within each parameter category, \ac{do} was sensitive to three parameters for optics (75\% of all optic parameters), eight for organic matter (67\%), 16 for phytoplankton (73\%), one for temperature (50\%), and 10 for zooplankton (91\%). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, average sensitivity was $L1 = $ $9.2\times 10^{-3}$ with values ranging from $L1 = $ $8.34\times 10^{-8}$ for \textit{QminP} (phytoplankton) to $0.05$ for \textit{umax} (phytoplankton). Within categories (excluding temperature with one sensitive parameter), sensitivity ranged from $4.39\times 10^{-5}$ (\textit{astarOMA}) to $7.51\times 10^{-4}$ (\textit{astar490}) for optics, $4.17\times 10^{-4}$ (\textit{KNH4}) to $6.15\times 10^{-3}$ (\textit{KG1}) for organic matter, $8.34\times 10^{-8}$ (\textit{QminP}) to $0.05$ (\textit{umax}) for phytoplankton, and $3.69\times 10^{-5}$ (\textit{ZQp}) to $0.05$ (\textit{ZKa}) for zooplankton (\cref{tab:dosens}).  Average sensitivity values in each category were $L1 = $ $2.81\times 10^{-4}$ for optics, $2.17\times 10^{-3}$ for organic matter, $0.02$ for temperature, $0.01$ for phytoplankton, and $0.01$ for zooplankton.

Local sensitivity analyses for the additional state variables (ammonium, \ac{chla}, irradiance, nitrate, \ac{pom}, \ac{dom}, and phosphorus) had similar results as \ac{do} with some exceptions (\cref{fig:sensalltile,tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens}).  All variables were sensitive to the same parameters as \ac{do} (38 of 51 evaluated), although average sensitivity differed between variables.  Average $L1$ ranged from $0.02$ for irradiance (\cref{tab:irrsens}) to $0.71$ for \ac{dom} (\cref{tab:om2sens}).  All average sensitivity values for the state variables were higher than the average for \ac{do} ($L1$ = $9.2\times 10^{-3}$).  For each variable, $L1$ ranged from $2.24\times 10^{-6}$ (\textit{QminP}) to $8.49$ (\textit{mA}) for ammonium (\cref{tab:nh4sens}), $1.38\times 10^{-6}$ (\textit{QminP}) to $13.94$ (\textit{mA}) for \ac{chla} (\cref{tab:chlsens}), $1.92\times 10^{-7}$ (\textit{QminP}) to $0.13$ (\textit{ZKa}) for irradiance (\cref{tab:irrsens}), $6.67\times 10^{-7}$ (\textit{QminP}) to $8.49$ (\textit{umax}) for nitrate (\cref{tab:no3sens}), $6.41\times 10^{-5}$ (\textit{KNH4}) to $7.22$ (\textit{mA}) for \ac{pom} (\cref{tab:om1sens}),  $7.41\times 10^{-5}$ (\textit{KNH4}) to $14.25$ (\textit{mA}) for \ac{dom} (\cref{tab:om2sens}), and $8.21\times 10^{-7}$ (\textit{QminP}) to $1.47$ (\textit{ZKa}) for phosphate (\cref{tab:po4sens}).  For the parameter categories, ammonium was most sensitive to phytoplankton parameters (average $L1$ = $0.8$ across all parameters in the category), \ac{chla} to phytoplankton ($L1$ = $1.14$), irradiance to zooplankton ($L1$ = $0.03$), nitrate to zooplankton ($L1$ = $1.06$), \ac{pom} to temperature ($L1$ = $0.86$), \ac{dom} to temperature ($L1$ = $1.48$), and phosphate to zooplankton ($L1$ = $0.31$).  Finally, average sensitivity between parameter categories independent of the state variables ranged from $8.38\times 10^{-3}$ for optics (average $L1$ across all variables) to $0.62$ for phytoplankton.

@@

Calibration of parameter values for each of the seven selection heuristics (\cref{tab:heurist1,tab:heurist2,tab:heurist3})  produced model output that varied in goodness of fit and computational effort required to identify optimal values (\cref{tab:calib}).  As expected, the number of required iterations of the optimization algorithm to minimize \ac{rmse} of observed \ac{do} and model output varied in proportion to the number of parameters in a subset.  For example, only three iterations were required to optimize the single parameter selected for the temperature category, whereas 136 iterations were required for the eight parameters selected independent of category.  No association was observed between collinearity ($\gamma$) and number of iterations or calibration.  Of the seven parameter subsets, selection independent of parameter category and equally within each category produced the largest reduction in \ac{rmse} values from the default output (starting \ac{rmse} $35.17$ mmol O$_2$ m$^{-3}$, reduction to $35.15$, $35.16$, respectively), whereas the optimal values for parameters in the Optics category had the smallest reduction ($35.17$).  Lower \ac{rmse} was associated with higher average sensitivity of a parameter subset.  Parameters selected only in the optics category had the highest \ac{rmse} and lowest average L1 ($2.81\times 10^{-4}$, three parameters), whereas parameters selected independent of category had the lowest \ac{rmse} and highest average L1 ($0.03$, eight parameters, \cref{tab:calib}).

The estimated areal extent of bottom-water hypoxia using default parameter values for the 3-dimensional model peaked on August, 28\textsuperscript{th} at $1.19\times 10^{4}$ km$^2$ (\cref{fig:areachg}a, August mean area 8988 $\pm$ 551 km$^2$, \cref{fig:areachg}b). Increases in the top sensitive parameters for \ac{do} in each category (\cref{tab:dosens}) increased the estimated areal extent of hypoxia (Figs. \ref{fig:areachg}c, \ref{fig:areachg}e), with the exception of the parameter for optimum temperature of growth (\textit{Tref(nospA+nospZ)$_{p1}$}), which caused a decrease (mean difference from July default area -1070 $\pm$ 1161 km$^2$, \cref{fig:areachg}d).  The largest increase in area was caused by an increase in the maximum growth rate of phytoplankton (\textit{umax}), with the greatest increase occurring in July (monthly mean increase of 4109 $\pm$ 1085 km$^2$, \cref{fig:areachg}d).  Overall, changes in hypoxia extent were largest for parameters with large sensitivity values (i.e., L1 = $0.05$ for \textit{ZKa}, L1 = $0.05$ for \textit{umax}).

@@

@@

{\tiny

@@

~~initial slope of photosynthesis v irradiance&\textit{alpha}&10$^{-16}$ cm$^2$ s quanta$^{-1}$ d$^{-1}$&$8.42\times 10^{-17}$&$1.26\times 10^{-16}$\tabularnewline

@@

~~phytoplankton threshold for grazing&\textit{Athresh}&10$^7$ cells m$^{-3}$&$1.72\times 10^{8}$&$2.58\times 10^{8}$\tabularnewline

@@

~~phytoplankton carbon/cell&\textit{Qc}&10$^{-7}$ mmol C cell$^{-1}$&$1.35\times 10^{-6}$&$2.03\times 10^{-6}$\tabularnewline
~~minimum N cell-quota&\textit{QminN}&10$^{-9}$ mmol N cell$^{-1}$&$6.08\times 10^{-9}$&$9.12\times 10^{-9}$\tabularnewline
~~minimum P cell-quota&\textit{QminP}&10$^{-9}$ mmol P cell$^{-1}$&$6.19\times 10^{-10}$&$9.29\times 10^{-10}$\tabularnewline

@@

~~N-uptake rate measured at umax&\textit{vmaxN}&10$^{-8}$ mmol cell$^{-1}$ d$^{-1}$&$4.1\times 10^{-8}$&$6.15\times 10^{-8}$\tabularnewline
~~P-uptake rate measured at umax&\textit{vmaxP}&10$^{-8}$ mmol cell$^{-1}$ d$^{-1}$&$2.68\times 10^{-8}$&$4.02\times 10^{-8}$\tabularnewline
~~Si-uptake rate measured at umax&\textit{vmaxSi}&10$^{-8}$ mmol cell$^{-1}$ d$^{-1}$&$4.1\times 10^{-8}$&$6.15\times 10^{-8}$\tabularnewline
~~phytoplankton volume/cell&\textit{volcell}&$\mu$m$^3$&$3.37\times 10^{4}$&$5.05\times 10^{4}$\tabularnewline

@@

~~half saturation coefficient for grazing&\textit{ZKa}&$\mu$m$^3$ m$^{-3}$&$1.12\times 10^{12}$&$1.68\times 10^{12}$\tabularnewline
~~quadratic mortality constant&\textit{Zm}&m$^6$ ind$^{-2}$ d$^{-1}$&$7.2\times 10^{-4}$&$0$\tabularnewline
~~zooplankton carbon/individual&\textit{ZQc}&mmol C ind$^{-1}$&$3.13\times 10^{-4}$&$4.7\times 10^{-4}$\tabularnewline
~~zooplankton nitrogen/individual&\textit{ZQn}&mmol N ind$^{-1}$&$6.95\times 10^{-5}$&$1.04\times 10^{-4}$\tabularnewline
~~zooplankton phosphorus/individual&\textit{ZQp}&mmol P ind$^{-1}$&$3.77\times 10^{-6}$&$5.66\times 10^{-6}$\tabularnewline

@@

~~maximum growth rate of zooplankton&\textit{Zumax}&$\mu$m$^3$ ind$^{-1}$ d$^{-1}$&$9.45\times 10^{7}$&$1.42\times 10^{8}$\tabularnewline
~~zooplankton volume/individual&\textit{Zvolcell}&$\mu$m$^3$ ind$^{-1}$&$2.98\times 10^{7}$&$4.47\times 10^{7}$\tabularnewline

@@

~~Chla specific absorption at 490 nm&\textit{astar490}&$7.51\times 10^{-4}$\tabularnewline
~~OMZ specific absorption at 490 nm&\textit{astarOMZ}&$4.92\times 10^{-5}$\tabularnewline
~~OMA specific absorption at 490 nm&\textit{astarOMA}&$4.39\times 10^{-5}$\tabularnewline

@@

~~turnover rate for OM1A and OM1Z&\textit{KG1}&$6.15\times 10^{-3}$\tabularnewline
~~turnover rate for OM2A and OM2Z&\textit{KG2}&$3.14\times 10^{-3}$\tabularnewline
~~O2 concentration that inhibits denitrification&\textit{KstarO2}&$3.04\times 10^{-3}$\tabularnewline
~~decay rate of CDOM&\textit{KGcdom}&$2.98\times 10^{-3}$\tabularnewline
~~half-saturation concentration for O2 utilization&\textit{KO2}&$5.85\times 10^{-4}$\tabularnewline
~~half-saturation concentration for denitrification&\textit{KNO3}&$5.8\times 10^{-4}$\tabularnewline
~~maximum rate of nitrification per day&\textit{nitmax}&$4.99\times 10^{-4}$\tabularnewline
~~NH4 rate constant for nitrification&\textit{KNH4}&$4.17\times 10^{-4}$\tabularnewline

@@

~~phytoplankton growth respiration coefficient&\textit{respg}&$8.36\times 10^{-3}$\tabularnewline
~~N-uptake rate measured at umax&\textit{vmaxN}&$8.12\times 10^{-3}$\tabularnewline
~~phytoplankton basal respiration coefficient&\textit{respb}&$6.94\times 10^{-3}$\tabularnewline
~~phytoplankton threshold for grazing&\textit{Athresh}&$4.57\times 10^{-3}$\tabularnewline
~~minimum N cell-quota&\textit{QminN}&$4.32\times 10^{-3}$\tabularnewline
~~P-uptake rate measured at umax&\textit{vmaxP}&$4.27\times 10^{-3}$\tabularnewline
~~coefficient for non-limiting nutrient&\textit{aN}&$4.23\times 10^{-3}$\tabularnewline
~~phytoplankton volume/cell&\textit{volcell}&$4.13\times 10^{-3}$\tabularnewline
~~half-saturation constant for P&\textit{Kp}&$2.9\times 10^{-3}$\tabularnewline
~~half-saturation constant for N&\textit{Kn}&$2.77\times 10^{-4}$\tabularnewline
~~minimum P cell-quota&\textit{QminP}&$8.34\times 10^{-8}$\tabularnewline

@@

~~proportion of phytoplankton lost to sloppy feeding&\textit{Zslop}&$7.78\times 10^{-3}$\tabularnewline
~~zooplankton growth-dependent respiration factor&\textit{Zrespg}&$5.32\times 10^{-3}$\tabularnewline
~~zooplankton biomass-dependent respiration factor&\textit{Zrespb}&$2.96\times 10^{-3}$\tabularnewline
~~zooplankton carbon/individual&\textit{ZQc}&$9.38\times 10^{-5}$\tabularnewline
~~zooplankton phosphorus/individual&\textit{ZQp}&$3.69\times 10^{-5}$\tabularnewline

@@

~~4&\scriptsize{\textit{KG1}}&$6.15\times 10^{-3}$&$1$\textsubscript{OM}&$3.85$\tabularnewline
~~5&\scriptsize{\textit{astar490}}&$7.51\times 10^{-4}$&$1$\textsubscript{O}&$3.89$\tabularnewline

@@

~~8&\scriptsize{\textit{KG1}}&$3.96\times 10^{-3}$&$2$\textsubscript{OM}&$14.65$\tabularnewline
~~9&\scriptsize{\textit{astarOMA}}&$1.47\times 10^{-3}$&$2$\textsubscript{O}&$14.72$\tabularnewline

@@

~~OMA specific absorption at 490 nm&\textit{astarOMA}&$1.63\times 10^{-3}$\tabularnewline

>>>>

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Wed Aug 16 18:18:02 2017 -0500, \href{https://github.com/fawda123/identifiability/commit/b081f900ad818adbbf496e0b8888820e38bda065}{b081f900ad818adbbf496e0b8888820e38bda065}\newline Acronyms: \acfi{chla}; \acfi{cgem}; \acfi{do}; \acfi{dom}; \acfi{gom}; \acfi{lcs}; \acfi{marb}; \acfi{pom}; \acfi{rmse}; \acfi{zerod}}

@@

\begin{abstract}
\noindent Local sensitivity analyses and identifiable parameter subsets were used to describe numerical constraints of a hypoxia model for bottom waters of the northern Gulf of Mexico.  The sensitivity of state variables differed considerably with parameter changes, although most variables were responsive to changes in parameters that influenced planktonic growth rates and less sensitive to physical or chemical parameters.  Variation in sensitivity had a direct correspondence with identifiability, such that only small subsets of the complete parameter set had unique effects on the model output. Selecting parameters by decreasing sensitivity demonstrated that only eight of 51 total parameters had a sufficiently unique effect on model output for accurate calibration.  As a result, parameter selection heuristics were used to identify parameters for model calibration that depended on combined effects on output, relative sensitivity of each parameter, and ecological categories for the biogeochemical equations. The calibrated \ac{zerod} unit of the hypoxia model had improved fit to the observed data if sensitive phytoplankton parameters were included in an identifiable subset.  Extension of results to a three-dimensional grid of the Gulf of Mexico showed that sensitive parameters for the \ac{zerod} model translated to non-trivial changes in the areal estimates of hypoxia.
\end{abstract}

@@

Selected parameter subsets from the seven heuristics were used to calibrate the \ac{zerod} model.  This analysis demonstrated that different parameter sets can produce model output with differing goodness of fit and calibration effort given
\begin{inparaenum}[1\upshape)]
\item differences in variable sensitivity to parameters in each subset, 
\item differences in the parameters selected from different categories (heuristic difference), and 
\item differences in identifiability values.
\end{inparaenum}
Generally, this analysis was a proof of concept that the selection heuristics can lead to an `optimal' model by maximizing the balance between sensitivity and identifiability.

Estimates of oxygen production from closed bottle experiments were used to calibrate the \ac{zerod} model.  Surface water was collected from a fixed location in Pensacola Bay, Florida (30.35$^{\circ}$ N, 87.20$^{\circ}$ W) on September 25\textsuperscript{th}, 2013 (full details in \citetalias{MurrellIR}).  Water samples were placed into closed 100 mL glass bottles at the time of collection and transported back to the lab for treatments.  Concentrations of \ac{do} and nutrients (NH$_4^+$, NO$_2^-$, NO$_3^-$, PO$_4^{3-}$, and SiO$_3^{2-}$) were extracted from the samples within one hour of collection. Oxygen concentrations were established using Winkler reagents \citep{Parsons85} and titration using a MetrOhm Titrando with thiosulfate titrant calibrated against an iodate standard and electrochemical endpoint detection.  Ammonium was analyzed using a fluorometric method \citep{Holmes99}; other nutrients were analyzed using either a Thermo Fisher Aquakem 200 discrete analyzer or an Astoria-Pacific continuous flow analyzer using standard colorimetric methods \citepalias{APHA05}.

Oxygen production after an approximate 12 hour period (morning to evening) was estimated within the bottles using light and dark treaments.  All treatments were exposed to ambient outdoor conditions in a water bath with continuous flow at the laboratory.  The dark treatments used opaque black bottles to ensure no exposure of ambient light to the samples.  The exposure experiments began at 7am and were concluded at 7pm, after which the concentrations of \ac{do} in each bottle were measured.  Each treatment had four replicates and HOBO\textsuperscript{\textregistered} pendants were used to measure continuous temperature and light.

Each parameter subset in \cref{tab:heurist1,tab:heurist2,tab:heurist3} for \ac{do} was calibrated to the ending oxygen concentration for each of the four replicates of each treatment. The intial nutrient and oxygen concentrations were used as starting conditions for FishTank.  The corresponding continuous temperature and light measurements were also used as forcing conditions.  Parameter values were searched using the optim function in R that implemented the limited memory modification of the BFGS quasi-Newton method \citep{Byrd95,Nocedal06,RDCT17}.  This optimization scheme searches for values within the user-defined lower and upper bounds of each parameter.  We set the algorithm to search within a range of values for a subset that was chosen to best reflect expectations given professional judgment and reasonable bounds from the literature.  While we recognize that `actual' parameter values could possibly extend beyond these ranges, constraining the values was expected to produce output that was more interpretable within the context of the sensitivity functions defined above.  Moreover, precise limits on each parameter were less of a concern for the current analysis as our primary goal was to demonstrate differences in fit as a function of sensitivity and identifiability.  The optimal values for each parameter were identified based on convergence of the model results as a minimization of the \ac{rmse} between the observed and modelled data.

@@

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in 38 of the 51 (75\% of total) parameters that were evaluated in FishTank (\cref{fig:sensalltile}, \cref{tab:dosens}). Within each parameter category, \ac{do} was sensitive to three parameters for optics (75\% of all optic parameters), eight for organic matter (67\%), 16 for phytoplankton (73\%), one for temperature (50\%), and 10 for zooplankton (91\%). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, average sensitivity was $L1 = $ $9.2e^{-3}$ with values ranging from $L1 = $ $8.34e^{-8}$ for \textit{QminP} (phytoplankton) to $0.05$ for \textit{umax} (phytoplankton). Within categories (excluding temperature with one sensitive parameter), sensitivity ranged from $4.39e^{-5}$ (\textit{astarOMA}) to $7.51e^{-4}$ (\textit{astar490}) for optics, $4.17e^{-4}$ (\textit{KNH4}) to $6.15e^{-3}$ (\textit{KG1}) for organic matter, $8.34e^{-8}$ (\textit{QminP}) to $0.05$ (\textit{umax}) for phytoplankton, and $3.69e^{-5}$ (\textit{ZQp}) to $0.05$ (\textit{ZKa}) for zooplankton (\cref{tab:dosens}).  Average sensitivity values in each category were $L1 = $ $2.81e^{-4}$ for optics, $2.17e^{-3}$ for organic matter, $0.02$ for temperature, $0.01$ for phytoplankton, and $0.01$ for zooplankton.

Local sensitivity analyses for the additional state variables (ammonium, \ac{chla}, irradiance, nitrate, \ac{pom}, \ac{dom}, and phosphorus) had similar results as \ac{do} with some exceptions (\cref{fig:sensalltile,tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens}).  All variables were sensitive to the same parameters as \ac{do} (38 of 51 evaluated), although average sensitivity differed between variables.  Average $L1$ ranged from $0.02$ for irradiance (\cref{tab:irrsens}) to $0.71$ for \ac{dom} (\cref{tab:om2sens}).  All average sensitivity values for the state variables were higher than the average for \ac{do} ($L1$ = $9.2e^{-3}$).  For each variable, $L1$ ranged from $2.24e^{-6}$ (\textit{QminP}) to $8.49$ (\textit{mA}) for ammonium (\cref{tab:nh4sens}), $1.38e^{-6}$ (\textit{QminP}) to $13.94$ (\textit{mA}) for \ac{chla} (\cref{tab:chlsens}), $1.92e^{-7}$ (\textit{QminP}) to $0.13$ (\textit{ZKa}) for irradiance (\cref{tab:irrsens}), $6.67e^{-7}$ (\textit{QminP}) to $8.49$ (\textit{umax}) for nitrate (\cref{tab:no3sens}), $6.41e^{-5}$ (\textit{KNH4}) to $7.22$ (\textit{mA}) for \ac{pom} (\cref{tab:om1sens}),  $7.41e^{-5}$ (\textit{KNH4}) to $14.25$ (\textit{mA}) for \ac{dom} (\cref{tab:om2sens}), and $8.21e^{-7}$ (\textit{QminP}) to $1.47$ (\textit{ZKa}) for phosphate (\cref{tab:po4sens}).  For the parameter categories, ammonium was most sensitive to phytoplankton parameters (average $L1$ = $0.8$ across all parameters in the category), \ac{chla} to phytoplankton ($L1$ = $1.14$), irradiance to zooplankton ($L1$ = $0.03$), nitrate to zooplankton ($L1$ = $1.06$), \ac{pom} to temperature ($L1$ = $0.86$), \ac{dom} to temperature ($L1$ = $1.48$), and phosphate to zooplankton ($L1$ = $0.31$).  Finally, average sensitivity between parameter categories independent of the state variables ranged from $8.38e^{-3}$ for optics (average $L1$ across all variables) to $0.62$ for phytoplankton.

@@

Calibration of parameter values for each of the seven selection heuristics (\cref{tab:heurist1,tab:heurist2,tab:heurist3}) produced model output that varied in goodness of fit and computational effort required to identify optimal values (\cref{tab:calib}).  As expected, the number of required iterations of the optimization algorithm to minimize \ac{rmse} of observed \ac{do} and model output varied in proportion to the number of parameters in a subset.  For example, only 12 iterations were required to optimize the single parameter selected for the temperature category, whereas 615 iterations were required for the seven parameters selected for phytoplankton (\cref{fig:calex}).  No association was observed between collinearity ($\gamma$) and number of iterations or calibration.  Of the seven parameter subsets, the optimal phytoplankton parameters had the largest reduction in \ac{rmse} values from the default output (starting \ac{rmse} $21.14$ mmol O$_2$ m$^{-3}$, reduction to $5.81$).  The identifiable parameters selected equally within categories (third heuristic) and independent of category (second heuristic) produced the second and third greatest reduction in \ac{rmse} ($19.09$, and $19.45$, respectively), although much less than that observed for phytoplankton parametrs. Overall, calibration to most of the parameter subsets had a small effect on \ac{rmse}.  However, the magnitude of parameter changes for calibration varied up to several orders of magnitude from the initial values (\cref{tab:calibprm}).  The percent change in parameters from the intial values varied from no change (Optics parameters \textit{astarOMZ} and ) to 900\% (\textit{KO2}).  Many calibrated parameter values included those at the lower or upper bounds used for the optimization routine (as for phytoplankton, \cref{fig:calex}).  There was no relationship between magnitude of parameter changes within a subset and overall reduction of \ac{rmse}.

The estimated areal extent of bottom-water hypoxia using default parameter values for the 3-dimensional model peaked on August, 28\textsuperscript{th} at $1.19e^{4}$ km$^2$ (\cref{fig:areachg}a, August mean area 8988 $\pm$ 551 km$^2$, \cref{fig:areachg}b). Increases in the top sensitive parameters for \ac{do} in each category (\cref{tab:dosens}) increased the estimated areal extent of hypoxia (Figs. \ref{fig:areachg}c, \ref{fig:areachg}e), with the exception of the parameter for optimum temperature of growth (\textit{Tref(nospA+nospZ)$_{p1}$}), which caused a decrease (mean difference from July default area -1070 $\pm$ 1161 km$^2$, \cref{fig:areachg}d).  The largest increase in area was caused by an increase in the maximum growth rate of phytoplankton (\textit{umax}), with the greatest increase occurring in July (monthly mean increase of 4109 $\pm$ 1085 km$^2$, \cref{fig:areachg}d).  Overall, changes in hypoxia extent were largest for parameters with large sensitivity values (i.e., L1 = $0.05$ for \textit{ZKa}, L1 = $0.05$ for \textit{umax}).

@@

% for abbreviated journal names

@@

@@

~~initial slope of photosynthesis v irradiance&\textit{alpha}&10$^{-16}$ cm$^2$ s quanta$^{-1}$ d$^{-1}$&$8.42e^{-17}$&$1.26e^{-16}$\tabularnewline

@@

~~phytoplankton threshold for grazing&\textit{Athresh}&10$^7$ cells m$^{-3}$&$1.72e^{8}$&$2.58e^{8}$\tabularnewline

@@

~~phytoplankton carbon/cell&\textit{Qc}&10$^{-7}$ mmol C cell$^{-1}$&$1.35e^{-6}$&$2.03e^{-6}$\tabularnewline
~~minimum N cell-quota&\textit{QminN}&10$^{-9}$ mmol N cell$^{-1}$&$6.08e^{-9}$&$9.12e^{-9}$\tabularnewline
~~minimum P cell-quota&\textit{QminP}&10$^{-9}$ mmol P cell$^{-1}$&$6.19e^{-10}$&$9.29e^{-10}$\tabularnewline

@@

~~N-uptake rate measured at umax&\textit{vmaxN}&10$^{-8}$ mmol cell$^{-1}$ d$^{-1}$&$4.1e^{-8}$&$6.15e^{-8}$\tabularnewline
~~P-uptake rate measured at umax&\textit{vmaxP}&10$^{-8}$ mmol cell$^{-1}$ d$^{-1}$&$2.68e^{-8}$&$4.02e^{-8}$\tabularnewline
~~Si-uptake rate measured at umax&\textit{vmaxSi}&10$^{-8}$ mmol cell$^{-1}$ d$^{-1}$&$4.1e^{-8}$&$6.15e^{-8}$\tabularnewline
~~phytoplankton volume/cell&\textit{volcell}&$\mu$m$^3$&$3.37e^{4}$&$5.05e^{4}$\tabularnewline

@@

~~half saturation coefficient for grazing&\textit{ZKa}&$\mu$m$^3$ m$^{-3}$&$1.12e^{12}$&$1.68e^{12}$\tabularnewline
~~quadratic mortality constant&\textit{Zm}&m$^6$ ind$^{-2}$ d$^{-1}$&$7.2e^{-4}$&$0$\tabularnewline
~~zooplankton carbon/individual&\textit{ZQc}&mmol C ind$^{-1}$&$3.13e^{-4}$&$4.7e^{-4}$\tabularnewline
~~zooplankton nitrogen/individual&\textit{ZQn}&mmol N ind$^{-1}$&$6.95e^{-5}$&$1.04e^{-4}$\tabularnewline
~~zooplankton phosphorus/individual&\textit{ZQp}&mmol P ind$^{-1}$&$3.77e^{-6}$&$5.66e^{-6}$\tabularnewline

@@

~~maximum growth rate of zooplankton&\textit{Zumax}&$\mu$m$^3$ ind$^{-1}$ d$^{-1}$&$9.45e^{7}$&$1.42e^{8}$\tabularnewline
~~zooplankton volume/individual&\textit{Zvolcell}&$\mu$m$^3$ ind$^{-1}$&$2.98e^{7}$&$4.47e^{7}$\tabularnewline

@@

~~Chla specific absorption at 490 nm&\textit{astar490}&$7.51e^{-4}$\tabularnewline
~~OMZ specific absorption at 490 nm&\textit{astarOMZ}&$4.92e^{-5}$\tabularnewline
~~OMA specific absorption at 490 nm&\textit{astarOMA}&$4.39e^{-5}$\tabularnewline

@@

~~turnover rate for OM1A and OM1Z&\textit{KG1}&$6.15e^{-3}$\tabularnewline
~~turnover rate for OM2A and OM2Z&\textit{KG2}&$3.14e^{-3}$\tabularnewline
~~O2 concentration that inhibits denitrification&\textit{KstarO2}&$3.04e^{-3}$\tabularnewline
~~decay rate of CDOM&\textit{KGcdom}&$2.98e^{-3}$\tabularnewline
~~half-saturation concentration for O2 utilization&\textit{KO2}&$5.85e^{-4}$\tabularnewline
~~half-saturation concentration for denitrification&\textit{KNO3}&$5.8e^{-4}$\tabularnewline
~~maximum rate of nitrification per day&\textit{nitmax}&$4.99e^{-4}$\tabularnewline
~~NH4 rate constant for nitrification&\textit{KNH4}&$4.17e^{-4}$\tabularnewline

@@

~~phytoplankton growth respiration coefficient&\textit{respg}&$8.36e^{-3}$\tabularnewline
~~N-uptake rate measured at umax&\textit{vmaxN}&$8.12e^{-3}$\tabularnewline
~~phytoplankton basal respiration coefficient&\textit{respb}&$6.94e^{-3}$\tabularnewline
~~phytoplankton threshold for grazing&\textit{Athresh}&$4.57e^{-3}$\tabularnewline
~~minimum N cell-quota&\textit{QminN}&$4.32e^{-3}$\tabularnewline
~~P-uptake rate measured at umax&\textit{vmaxP}&$4.27e^{-3}$\tabularnewline
~~coefficient for non-limiting nutrient&\textit{aN}&$4.23e^{-3}$\tabularnewline
~~phytoplankton volume/cell&\textit{volcell}&$4.13e^{-3}$\tabularnewline
~~half-saturation constant for P&\textit{Kp}&$2.9e^{-3}$\tabularnewline
~~half-saturation constant for N&\textit{Kn}&$2.77e^{-4}$\tabularnewline
~~minimum P cell-quota&\textit{QminP}&$8.34e^{-8}$\tabularnewline

@@

~~proportion of phytoplankton lost to sloppy feeding&\textit{Zslop}&$7.78e^{-3}$\tabularnewline
~~zooplankton growth-dependent respiration factor&\textit{Zrespg}&$5.32e^{-3}$\tabularnewline
~~zooplankton biomass-dependent respiration factor&\textit{Zrespb}&$2.96e^{-3}$\tabularnewline
~~zooplankton carbon/individual&\textit{ZQc}&$9.38e^{-5}$\tabularnewline
~~zooplankton phosphorus/individual&\textit{ZQp}&$3.69e^{-5}$\tabularnewline

@@

~~4&\scriptsize{\textit{KG1}}&$6.15e^{-3}$&$1$\textsubscript{OM}&$3.85$\tabularnewline
~~5&\scriptsize{\textit{astar490}}&$7.51e^{-4}$&$1$\textsubscript{O}&$3.89$\tabularnewline

@@

~~8&\scriptsize{\textit{KG1}}&$3.96e^{-3}$&$2$\textsubscript{OM}&$14.65$\tabularnewline
~~9&\scriptsize{\textit{astarOMA}}&$1.47e^{-3}$&$2$\textsubscript{O}&$14.72$\tabularnewline

@@

~~OMA specific absorption at 490 nm&\textit{astarOMA}&$1.63e^{-3}$\tabularnewline
~~OMZ specific absorption at 490 nm&\textit{astarOMZ}&$1.5e^{-3}$\tabularnewline
